Stateful Forward Reasoning

### 2025-01-30 paper.tex
wip update

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Contribution Title}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%
\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath,amssymb}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
\usepackage{pgfplots}

\newtheorem{lem}{Lemma}

\setminted[lean4]{extrakeywords={aesop cases add aesop? intro simp simp_all only split apply on_goal next rename_i safe unsafe norm constructors forward destruct norm_num done add_aesop_rules rfl subst ext}}
\newmintinline[lean]{lean4}{bgcolor={},ignorelexererrors=true}
\newminted[leancode]{lean4}{bgcolor={},ignorelexererrors=true,fontsize=\footnotesize,autogobble}
\BeforeBeginEnvironment{leancode}{\begin{tcolorbox}}
\AfterEndEnvironment{leancode}{\end{tcolorbox}}
\usemintedstyle{xcode}

\newcommand{\xcom}[1]{{\color{cyan}{Xavier: #1}} }
\newcommand{\jcom}[1]{{\color{orange}{Jannis: #1}} }
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Stateful Forward Reasoning}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{University of Munich (LMU), Munich, Germany\\
\email{\{xavier.genereux,jannis.limpberg\}@lmu.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
150--250 words.

\keywords{Forward Reasoning \and Proof Automation \and Theorem Proving.}
\end{abstract}
%
%
%

\section{Introduction}
\xcom{TODO}

\xcom{ Specify the problem (user defined rules)
 + deal with changes with local context}
\xcom{ Talk about aesop}
\xcom{ Do we outline the solution?}

An important contextualization is that the forward reasoning procedure in this paper takes place
within aesop's routine procedure. We will focus on examples that highlight this style of proof
while noting that in practice, a combination of both forwards and backwards reasoning are used 
\section{Preliminaries}

\xcom{Adapt for CADE.}

In what follows, we write $\Gamma \models T$ for a goal where $\Gamma$ represents the context and
$T$ the target. In a mathematical problem, $\Gamma$ is the list of hypotheses and variable from
which we wish to show $T$.

\paragraph{Forward Rules.}
Consider the following lemma :
\begin{lem}[\lean{le_trans}]
  Given a preorder on some type $\alpha$ and elements $a,b,c$ of type $\alpha$, the following holds :
  $$a \leq b \rightarrow b \leq c \rightarrow a \leq c$$
\end{lem}

From such lemmas, we will create what we call \textit{forward rules} that represent
the theorems on which we would like to reason in a forward manner.
The premises $\alpha$, $ \text{Preorder } \alpha$ , $a$, $b$, $c$, $a \leq b$ and
$b \leq c$ are separated between \textit{implied hypotheses} $\text{Preorder } \alpha$,
$a$, $b$ and $c$ and \textit{maximal hypotheses} $a \leq b$ and $b \leq c$.
Maximal hypotheses are also called \textit{input hypotheses}.

For the premises $(x_1 : T_1),(x_2 : T_2),\dots,(x_n : T_n)$,
the input hypotheses consist of the hypotheses for which there are no
$j > i$ such that $T_j$ depends on $x_i$.

\paragraph{Rules.}
We write $r : \Phi \to \Psi$ for a \textit{rule} $r$ with input
hypotheses $\Phi$ and output hypothesis $\Psi$.
For a given rule, we assign each maximal input hypothesis a unique index
between $1$ and $n$.
We call theses indices \textit{slots} of $r$ and define $\text{slots}(r)$ as
the set of slots of $r$.

Given a slot $i$ and hypothesis $h : T$ with $T \equiv_{[\sigma]} T_i$
(i.e. $\sigma$ is the substitution that results from unifying $T$ and $T_i$), 
we say that $h$ is \textit{suitable for slot $i$} and define $\text{sub}(h) ≔ \sigma$.

A \textit{shared variable} of $r$ is an input hypothesis $x_i$ such that there are
at least two input hypothesis types $T_j$, $T_k$ in which $x_i$ occurs.
A shared variable is necessarily a non-maximal input hypothesis.
We also write $?x_i$ to emphasise that we interpret $x_i$ as a variable
(and, in the code, as a metavariable).
The set $s_r(i)$ of \textit{shared variables of a slot} $i$ of $r$ are those
shared variables $?x$ that occur in $T_i$ and in at least one slot $j < i$.
If the rule in question is clear from the context, we write $s(i)$.

%The typical behavior is that running the procedure in a context containing suitable
%candidates for the maximal hypotheses will add the \textit{result} of the forward
%rule $a < c$ to the context.
%The result is also called the \textit{output hypothesis}.
%If we find matching hypotheses for all maximal input hypotheses in the local context,
%we can apply the forward rule, since each non-maximal input hypothesis is uniquely determined
%by maximal hypotheses.

A \textit{match} $M$ for $r$ is a partial map from slots of $r$ to hypotheses in
the context.
It must satisfy the following conditions:
\begin{itemize}
\item For each slot $i$ of $r$, if $M(i)$ is defined, then
    $M(i) \equiv_{[\sigma_i]} T_i$ with a substitution $\sigma_i$.
    The domain of this substitution is a subset of the non-maximal input hypotheses.
\item For all slots $i$ and $j$ of $r$ such that $M(i)$ and $M(j)$ are defined,
    $\sigma_i$ and $\sigma_j$ are compatible.
\item $M$ is downward-closed: if $M(i)$ is defined, then for any slot $j < i$ of
    $r$, $M(j)$ is also defined.

\end{itemize}
A match of $r$ is \textit{complete} if $M(i)$ is defined for every slot $i$ of $r$, and
\textit{incomplete} otherwise.
The \textit{level of a match} $M$, $lvl(M)$, is the maximal $i$ such that $M(i)$ 
is defined.
The \textit{substitution of a match} $M$, $sub(M)$, is the union of the
substitutions $\sigma_i$ for each $i$ such that $M_i$ is defined.
This union is defined since the $\sigma_i$ are pairwise compatible.
This means that $sub(M)$ contains only one assignement per variable.
The set of \textit{shared variables of a match} $M$ is $s(lvl(M))$ the
shared variables of $lvl(M)$.

An output hypothesis is \textit{redundant} if its type is already present in the context.
A goal is saturated with respect to a set of forward rules if all
output hypotheses returned by the procedure are redundant.

The following section introduces an algorithm to construct complete matches $M$ 
from the context $\Gamma$.

\section{Design}

We now focus on the problem of constructing complete matches.

\subsection{A naive design}

A straightforward way to solve this problem is to simply test all the hypotheses present in the
context against the registered rules.

A few obvious improvements are then available. We can index the rules in terms of one of its
hypothesis. This index can then be used to guarentee that at least one hypothesis in the context will
match with a slot of the rule. A heuristic is to take the last slot as it should be more specific
than the firsts slot of a rule.

If the last slot of a rule matches a hypothesis in the context we can then test
whether other hypotheses of the context can complete that rule.
This is done in two steps, first we check whether the hypotheses in the context match the slot.
If it does, we check if the substitution given by the unification is coherent with
the preceding matched hypotheses.
Using the same reasoning as above, we heuristically do this backwards in the slots.

This particular design is fast when the final hypothesis of a rule is not present in the context. 
However, it otherwise needs to test many potential hypothesis candidate which is quite slow.
Another weakness of this approach is that in the eventuality that a new hypothesis appears in the
goal, for example as the result of another rule being completed, all previous work done on one rule
must be repeated with this new hypothesis.

We call stateless this kind of approach to the problem as no previous work is saved.

\subsection{An improved, stateful design}

With the weaknesses of the last approach in mind, we now consider a
\textit{stateful} solution to the problem of finding complete matches.
The main idea is to save partial progress on complete matches.

This partial progress could be simply saved as trees with root the matched
hypothesis returned by the index.

\paragraph{Indexing Scheme.}
However, we decided to move towards an index that looks at all the hypotheses of 
the rule.
The idea is that we can better track and save information about
which hypotheses will potentially complete which rules.
This makes it particularly easy to maintain partial matches which, as we will see,
are classified in terms of the instantiation of their set of shared variables.
Another advantage is that we only need to look at each of the context's
hypotheses once.
This makes the procedure incremental, which is part of the stateful philosophy.

In other words, the indexing is now a discrimination tree that maps 
types $T$ to sets of pairs $(r,i)$ where $r : (x_1 : T_1) \dots (x_n : T_n) \to \Psi$
is a rule, $i$ is a slot of $r$ and $T$ is likely unifies with $T_i$.

Given hypotheses returned by this index, the next challenge comes
down to finding a complete match of a rule.
This means identifying compatible hypotheses within the hypotheses returned by the index.

\paragraph{Matches.}
Compatible hypotheses are hypotheses such that the instantiation of their
shared variables are the same.
For example, the hypotheses $h_1 : a \leq b$ and $h_2 : b \leq c$ are compatible
because they are instantiations of $x \leq y$ and $y \leq z$ where the
instantiation of their shared variable $\{y\}$, is the same $\{b\}$.
A complete match is a set of hypotheses such that each of these hypothesis corresponds
to exactly one slot and they are all compatible.
We proceed in the following way to find a complete match.

First, we order the slots such that each slot has at least one variable in common
with the previous slots. This can be proven to always be possible. 
\xcom{We can add the proof if we want to / have space.}

We then aim to construct matches iteratively : 
We consider each hypotheses one by one, assign them to their
respective slots and check whether there exist compatible matches
which they can extend.
The comptibility check means that all the shared variables of the considered 
hypotheses need to be compatible.
It would be costly to do this comparison everytime and indeed
we choose instead to maintain what we call \textit{variable maps} whose purpose
is to save the current partial matches.
Each variable map is associated to one shared variable, these are the
variables dictate compatibility.

\paragraph{Variable Maps.}
Let $x$ ba a shared variable, $t$ an expression that represents an
instantiation of that variable and $i$ an integer associated to a slot.
We define two variable maps $\mu_x$ and $\eta_x$ that respectively take the 
pair $(i,t)$ to $\mathbb{M}$ and $\mathbb{H}$ where 
\begin{itemize}
    \item $\mathbb{M}$ is the set that contains exactly those
    partial matches which already contain assignments for all
    slots up to $i$ and which instantiate $?x$ with $t$.
    \item $\mathbb{H}$ is the set of hypotheses which match $T_i$ while
    instantiating $?x$ with $t$.
\end{itemize}

Given a hypothesis $h : T$ that matches $T_i$ and $\sigma_h$,
the map sending the variable of $T_i$ to their specific instantiation for $h$.
The compatibility check that looks for suitable partial matches is simply computed by
$$\bigcap_{s(i)} \mu_x \left(i - 1, \sigma_h(x)\right).$$

Similarly, given a match $m$ and $\sigma_m$ the substitution map of the relevant 
variable in $m$ then we can check for compatible hypotheses that would extend $m$
by computing 
$$\bigcap_{s(lvl(M) + 1)} \eta_x \left(i + 1, \sigma_m(x)\right).$$

\paragraph{Example 1.}
\xcom{WIP}
To demonstrate the typical behavior of the algorithm we consider the
following example which has \textbf{Lemma 1} as a rule and the hypotheses
$(h_1 : 300 \leq 301)$ and $(h_2 : 301 \leq 302)$. \xcom{format this.}
We will consider the hypothesis $h_2$ first for demonstration purposes.

At the beginning, for the rule \textit{le\_trans}, we notice that the
only shared variable is ${y}$. Now the hypothesis $(h_2)$ unifies both
with slot 1 $x \leq y$ and slot 2 $y \leq z$.
This updates the map $\eta_y(y \mapsto 302, 1)$ with $\{h_2\}$ and 
$\eta_y(y \mapsto 301, 2)$ with $\{h_2\}$

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
    \node[xshift = 0] {$\eta_y$}
      child[xshift = -20] {node {1}
          child {node {$302$}
            child {node {$\{h_2\}$}}}
          child {node {$301$}
            child {node {$\{h_1\}$}}}}
      child[xshift = 20] {node {2}
          child {node {$301$}
            child {node {$\{h_2\}$}}}
          child {node {$300$}
            child {node {$\{h_1\}$}}}};
    \node[xshift = 90] {$\eta_y$}
        child[xshift = -20] {node {1}
            child {node {$302$}
                child {node {$\{h_2\}$}}}
            child {node {$301$}
                child {node {$\{h_1\}$}}}}
        child[xshift = 20] {node {2}
            child {node {$301$}
                child {node {$\{h_2\}$}}}
            child {node {$300$}
                child {node {$\{h_1\}$}}}};
    \node[xshift=200] {$\mu_y$}
      child {node[xshift=-20] {1}
          child {node {$302$}
            child {node {$\{\{1 \mapsto h_2\}\}$}}}
          child {node {$301$}
            child {node {$\{\{1 \mapsto h_1\}\}$}}}}
      child {node[xshift=20] {2}
          child {node {$301$}
            child {node[yshift=-2em] {$\{\{1 \mapsto h_1, 2 \mapsto h_2\}\}$}}}};
  \end{tikzpicture}

\paragraph{Example 2.}
We now consider another example with more than one shared variable.
Let $T_1, T_2$ be two proposition that both depend on the variable $x$ and $y$.
For simplicity, the rule considred will be $T_1(x,y) \to T_2(x,y) \to R$.
The hypotheses added to the context will be, in order,
$h_1 : T_2(a,a)$, $h_2 : T_1(a,b)$ and $h_3 : T_2(a,b)$.

\begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
    \node[xshift = 0] {$\eta_x$}
      child[xshift = -20] {node {1}
          child {node {$a$}
            child {node {$\{ \}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}}
      child[xshift = 20] {node {2}
          child {node {$a$}
            child {node {$\{h_1\}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}};
    \node[xshift = 90] {$\eta_y$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{ \}$}}}}
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            child {node {$b$}
                child {node {$\{ \}$}}}};
  \end{tikzpicture}

  At this point, there are no matches so there is nothing to extend.
  We add the next hypothesis in the queue to the variable maps.

  \begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
    \node[xshift = 0] {$\eta_x$}
      child[xshift = -20] {node {1}
          child {node {$a$}
            child {node {$\{h_2\}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}}
      child[xshift = 20] {node {2}
          child {node {$a$}
            child {node {$\{h_1\}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}};
    \node[xshift = 90] {$\eta_y$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{ h_2\}$}}}}
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            child {node {$b$}
                child {node {$\{ \}$}}}};
  \end{tikzpicture}

  As this hypothesis is in the first slot, we create the corresponding level 1 partial
  match $m_0 = {(h_2 : T_1(a, b))}$ for which we check for suitable hypotheses that could
  extend the match.
  Note that  $s(lvl(m_0) + 1) = s(2)$ which is the intersection of the variables
  appearing in the slots $i < 2$ with the variables appearing in slot $2$.
  Thus $s(2) = \{a,b\}$ and we compute :
  $$\bigcap_{x\in s(lvl(m_0) + 1)} \eta_x \left(i + 1, \sigma_{m_0}(x)\right)
    = \eta_x(2, a) \cap \eta_y(2, b) = \emptyset $$
  Since the set is empty, the match can not be currently extended.
  We turn to the remaining hypothesis in the queue and add it to the maps.

  \begin{tikzpicture}[outer sep=auto, level distance=10mm, scale=0.50]
    \node[xshift = 0] {$\eta_x$}
      child[xshift = -20] {node {1}
          child {node {$a$}
            child {node {$\{h_2\}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}}
      child[xshift = 20] {node {2}
          child {node {$a$}
            child {node {$\{h_1, h_3\}$}}}
          child {node {$b$}
            child {node {$\{ \}$}}}};
    \node[xshift = 90] {$\eta_y$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{h_2\}$}}}}
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            child {node {$b$}
                child {node {$\{h_3\}$}}}};
  \end{tikzpicture}

  We are now ready to look for potential matches that could be extended by $h_3$.
  In this particuliar case, since $h_3$ is in slot $2$ we look for $s(2) = \{a,b\}$.
  Then,
  $$\bigcap_{x\in s(2)} \mu_x \left(i - 1, \sigma_h(x)\right)
    = \mu_x(1, a) \cap \mu_y(1, b) = \{m_0\} $$.

  We have found two compatible hypotheses for our rule and we can then add the 
  result $R$ as hypothesis to the context.

\section{Implementation}

\xcom{Here are some notes about the implementation.}
The index used to match hypotheses with their matching rules are implemented as
discrimination trees. A discrimination tree is a trie-like data
structure that maps expressions to arbitrary data.
In \cite{}, it is shown to be good solution for the problem of finding
terms (here the types of hypotheses) that might unify with other terms
(here the types of the input hypotheses).
In Lean 4, discrimination trees are also used to index typeclass
instances and simplifier lemmas.

\xcom{Incrementality?}
In lean, the type of hypotheses can sometimes change while the hypothesis'
identifier remains the same. This means that one needs to track the types
of the hypotheses as well. We call this tracking \textit{goal diffs}, they
highlight the differences between old and updated goal states.
The reference to hypotheses which changing types get deleted from the maps
and the slot. We then add them again as a new hypothesis.

The maps described in the last section are all implemented as persistent
hash maps \xcom{Why did we not use discrimination trees? $+$
(pers. important cause we store related forward state for many goal.) }.

One of the more costly operation that we need to run multiple times is the
unification \xcom{defeq?} between the hypotheses and the types of the
input hypotheses associated to rules.
We implement a variant of Lean's unification process as follows :
The idea is that we wish to normalize the types as much as possible

\xcom{precompilation}

\section{Evaluation}

Analysing the both algorithms of both the stateless and stateful approach, we 
can draw the following conclusions about their relative performance :

\begin{itemize}
    \item The stateful algorithm maybe be expected to perform poorly when the
    context contains many hypotheses that almost complete rules.
    \item The stateless algorithm should perform decently well on average and 
    especially when only the last premises are absent; so the rules get never
    selected by the stateless index
\end{itemize}

\subsection{Results}
\xcom{TODO: Add a comment on the effect it had on Mathlib.}

We present here a number of results obtained by running synthetic tests.
By synthetic we mean that these tests we're constructed solely for the
purpose of demonstrating the different complexity behavior of both
algorithms by scaling some parameter. (e.g. the number of rules.)
The results of the tests are an average of multiple runs.

In \ref{fig:erase}, the step is set up as follows: 
Given some rule $r$ with $6$ \xcom{motivate 6} premises, 
we run the procedure in a context with 5 of the 6 premises present.
The numbers on the $x$ axis represent which hypothesis is missing
in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the algorithm investigate before
it stops.
At $0$, the rule are not selected by the index so the procedure stops.
The stateless algorithm gets progressively slower the further the removed
premises is located.

The position of the missing premise has little effect on the stateful
algorithm as the index selects the rule in all case; it only needs one
of the premise to be present.
The work the stateful algorithm has to do is thus close to constant
in this scenario.

\begin{figure}
% This is the erase without precompilation for 6 premises.
% Note I have a decent idea that would allow us to compare for multiple
% number of premises. 
% (Divide by the average of the stateful time to normalize.)
\begin{tikzpicture}
    \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        ylabel={Time in ms},
        xmin=0, xmax=6,
        ymin=50, ymax=65,
        xtick={0,1,2,3,4,5,6},
        ytick={50,55,60,65},
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
    ]

    \addplot[
        color=orange,
        mark=square,
        ]
        coordinates {
        (1,50.91)(2,53.14)(3,52.91)(4,51.22)(5,55.14)(6,56.60)
        };
        \addlegendentry{Stateless}

    \addplot[
        color=blue,
        mark=square,
        ]
        coordinates {
        (1,56.04)(2,56.43)(3,55.58)(4,55.26)(5,54.80)(6,60.61)
        };
        \addlegendentry{Stateful}

    \end{axis}
    \end{tikzpicture}
\end{figure}

\begin{figure}
    % This is the erase without precompilation for 6 premises.
    % Note I have a decent idea that would allow us to compare for multiple
    % number of premises. 
    % (Divide by the average of the stateful time to normalize.)
    \begin{tikzpicture}
        \begin{axis}[
            title={Scaled Depth Test},
            xlabel={Depth},
            ylabel={Time in ms},
            xmin=0, xmax=8,
            ymin=0.5, ymax=2,
            xtick={0,1,2,3,4,5,6,7,8},
            ytick={0.75,1,1.25,1.5,1.75},
            legend pos=north west,
            ymajorgrids=true,
            grid style=dashed,
        ]

        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.883600)(1, 1.039751)};
            \addlegendentry{Stateless}
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.804521)(1, 1.017893)(2, 1.022493)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.791528)(1, 1.015467)(2, 1.076267)(3, 1.159999)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.743256)(1, 0.966436)(2, 1.058876)(3, 1.133514)(4, 1.225561)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.770625)(1, 1.003777)(2, 1.109929)(3, 1.215171)(4, 1.305373)(5, 1.390667)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.759093)(1, 1.016667)(2, 1.148273)(3, 1.213138)(4, 1.342795)(5, 1.442271)(6, 1.514877)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.709965)(1, 0.971294)(2, 1.068745)(3, 1.179333)(4, 1.312798)(5, 1.424054)(6, 1.551690)(7, 1.628344)};
        \addplot[color=orange,mark=square,]
            coordinates {(0, 0.728860)(1, 1.001576)(2, 1.097472)(3, 1.219259)(4, 1.366683)(5, 1.444844)(6, 1.576972)(7, 1.668693)(8, 1.815399)};

        \end{axis}
        \end{tikzpicture}
    \end{figure}

\begin{figure}
    \label{fig:erase}
%erase with precompilation
    \begin{tikzpicture}
        \begin{axis}[
            title={Depth Test},
            xlabel={Depth},
            ylabel={Time in ms},
            xmin=0, xmax=6,
            ymin=3, ymax=7,
            xtick={0,1,2,3,4,5,6},
            ytick={3,4,5,6,7},
            legend pos=north west,
            ymajorgrids=true,
            grid style=dashed,
        ]

        \addplot[
            color=blue,
            mark=square,
            ]
            coordinates {
            (1, 3.715813)(2, 3.893547)(3, 3.856956)(4, 3.722684)(5, 3.714804)(6, 3.900508)
            };
            \addlegendentry{Stateful}

        \addplot[
            color=orange,
            mark=square,
            ]
            coordinates {
            (1, 2.993757)(2, 3.585232)(3, 3.880661)(4, 4.098640)(5, 4.322570)(6, 5.765420)
            };
            \addlegendentry{Stateless}

        \end{axis}
        \end{tikzpicture}
    \end{figure}

    \begin{figure}
        %Indep with precompilation
            \begin{tikzpicture}
                \begin{axis}[
                    title={Independance Test},
                    xlabel={Number of Rules},
                    ylabel={Time in ms},
                    xmin=0, xmax=32,
                    ymin=0, ymax=1700,
                    ymode=log,xmode=log,
                    log basis x=2,
                    log basis y=2,
                    legend pos=north west,
                    ymajorgrids=true,
                    grid style=dashed,
                ]

                \addplot[
                    color=blue,
                    mark=square,
                    ]
                    coordinates {
                    (1, 0.710963)(2, 1.266787)(4, 2.181111)(8, 4.220279)(16, 8.084744)(32, 15.683840)
                    };
                    \addlegendentry{Stateful}

                \addplot[
                    color=orange,
                    mark=square,
                    ]
                    coordinates {
                    (1, 0.872730)(2, 2.100940)(4, 7.245598)(8, 35.805959)(16, 222.283526)(32, 1600.978502)
                    };
                    \addlegendentry{Stateless}

                \end{axis}
                \end{tikzpicture}
            \end{figure}

            \begin{figure}
                %Casacade with precompilation
                    \begin{tikzpicture}
                        \begin{axis}[
                            title={Casacade Test},
                            xlabel={Number of Rules},
                            ylabel={Time in ms},
                            xmin=0, xmax=32,
                            ymin=0, ymax=850,
                            ymode=log,xmode=log,
                            log basis x=2,
                            log basis y=2,
                            legend pos=north west,
                            ymajorgrids=true,
                            grid style=dashed,
                        ]

                        \addplot[
                            color=blue,
                            mark=square,
                            ]
                            coordinates {
                            (1, 0.393011)(2, 0.730952)(4, 1.327736)(8, 3.007961)(16, 10.173689)(32, 38.892027)
                            };
                            \addlegendentry{Stateful}

                        \addplot[
                            color=orange,
                            mark=square,
                            ]
                            coordinates {
                            (1, 0.464973)(2, 0.759516)(4, 1.865191)(8, 7.585588)(16, 64.374740)(32, 813.336527)
                            };
                            \addlegendentry{Stateless}

                        \end{axis}
                        \end{tikzpicture}
                    \end{figure}

\section{Related Work}
\section{Conclusion}

\section{First Section}
\subsection{A Subsection Sample}
Please note that the first paragraph of a section or subsection is
not indented. The first paragraph that follows a table, figure,
equation etc. does not need an indent, either.
